Manuscript draft

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (**REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs. 2-3). This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**). Such observations may be attributable to the fact that temperate forests are predominantly secondary and of younger age than tropical forests (**REF**), so analyses comparing across latitudinal gradients without strict control of stand age tend to be comparing younger temperate stands with older tropical (and often boreal) forests ([Poulter et al. 2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)). Because C allocation varies with stand age (**REFS**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????**)]. By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.

[The fact that we detected no major trends in C allocation is quite signficant (even though analysis not designed specifically for this) (1) Discuss temperature senstivity of productivity vs respiration, CUE. While there's lots of support for this on shorter time scales, we don't find support at the annual scale. (2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]

>>>>

Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (**REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs. 2-3). This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**). Such observations may be attributable to the fact that temperate forests are predominantly secondary and of younger age than tropical forests (**REF**), so analyses comparing across latitudinal gradients without strict control of stand age tend to be comparing younger temperate stands with older tropical (and often boreal) forests ([Poulter et al. 2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)). Because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)]. By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.

We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig. 4), with no major trends in C allocatio (Fig. 2). This finding contradicts some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming. Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$). The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**). The latter understanding is driven by observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; @anderson-teixeira_carbon_2016] (**REFS(KAT)**), but our analysis suggests that these low values may more appropriately be attributed to the fact that these forests are old than to their climate. Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**). Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.  [There is some indication that $GPP$ may have a somewhat steeper relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig. 2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**). (**Helene, maybe comment on Larjavaara & Muller-Landau (2011)?**). ] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients.

(1) Discuss temperature senstivity of productivity vs respiration, CUE. While there's lots of support for this on shorter time scales, we don't find support at the annual scale. (2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]

We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes. In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients. There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006]. That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
